---
title: "Eatomics ssGSEA and Differential Enrichment Report"
date: '`r format(Sys.Date(), "%Y-%B-%d")`'
output: 
  html_document:
    theme: flatly
params:
  configuration: NA
  stats_proteinGroups: NA
  volcano_plot: NA
  boxPlotUp: NA
  boxPlotDown: NA
  UpRegul: NA
  DoRegul: NA
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Data description and quality control

The MaxQuant algrorithm identified `r params$stats_proteinGroups$NumFullProt` proteins across all samples. 
`r params$stats_proteinGroups$NumPotCon` potential contaminants, `r params$stats_proteinGroups$IdentifiedBySite` proteinsGroups only identified by site, and `r params$stats_proteinGroups$Reverse` proteins only identified in the reverse database were removed from the data set. The final dataset has `r params$stats_proteinGroups$Cleaned` rows. `r params$stats_proteinGroups$IntensityMetric` was used as intensity metric for single sample Gene Set Enrichment Analysis (ssGSEA) enrichment score calculation.
Missing values are imputed using the `r params$stats_proteinGroups$Imputation` method. The imputation methods are based on knn (k-nearest-neighbour) from the impute package [1], MinDet or QRLIC from the imputeLCMD package with default parameters [2] or a custom implementation of Perseus' sampling from a down-shifted Gaussian distribution (replaceMissingFromGaussian(), implemented by Matthias Ziehm) with default parameters of width = 0.3 and shift = 1.8.

## 2. Single samples Gene Set Enrichment Analysis (ssGSEA)

The calculation of ssGSEA (single-sample Gene Set Enrichment Analysis) scores is mainly a prerequisite to perform differential enrichment in step 4 and is adapted from Krug et al. [1] (https://github.com/broadinstitute/ssGSEA2.0). 

Each ssGSEA enrichment score represents the degree to which the genes in a particular gene set are coordinately up- or down-regulated within a single sample. 
Advantages of this approach include the flexible in- and exlcusion of samples without recalculation of differentially expressed proteins and presumably the possibility of reducing batch effects.

### 2.1 Configuration

Configuration of the ssGSEA algorithm implements the default parameters as stated at . 
The exponential weight employed in calculation of enrichment scores was set to `r params$configuration$weight`. Number of permutations were set to `r params$configuration$nperm`. The correlation type was set to `r params$configuration$correl.type`.
Gene sets whose overlap with the genes listed in the input GCT file are less than `r params$configuration$min.overlap` are excluded. 

### 2.2 Summary of ssGSEA results


## 3. Experimental setup of differential enrichment analysis

## 4. Results of differential enrichment analysis

```{r Experimental Setup, echo=F}
params$ExpSetup
```

### 4.1 Volcano plot
Differential enrichment volcano plot, x axis denoting the fold change and y axis the -log10 of P.value from the limma [`toptable` function](https://www.rdocumentation.org/packages/limma/versions/3.28.14/topics/toptable). P-values are multiple testing corrected by using the [Benjamini-Hochberg procedure](https://www.rdocumentation.org/packages/stats/versions/3.5.2/topics/p.adjust). 



```{r Limma Output, echo=F}

if(is.null(params$volcano_plot)){
  message("The volcano plot was not computed during the analysis and thus is not part of this report.\n")
} else {
  params$volcano_plot
}
```

### 4.2 Boxplots of selected enrichment scores

User selected pathways/gene sets showing higher enrichment scores.   

```{r Limma BoxplotUp, echo=F}

if(is.null(params$boxPlotUp)){
  message("The boxplot for enriched pathways (higher) was not computed during the analysis and thus is not part of this report.\n")
} else {
  params$boxPlotUp
}
```

User selected pathways/gene sets showing lower enrichment scores.   

```{r Limma BoxplotDown, echo=F}

if(is.null(params$boxPlotDown)){
  message("The boxplot for enriched pathways (lower) was not computed during the analysis and thus is not part of this report.\n")
} else {
  params$boxPlotDown
}
```

### 4.3 Table of pathways or gene sets with significant changes in enrichment

To see the full list of terms, jump to [appendix enrichments table (up)](#appendix-Up) and [appendix enrichments table (down)](#appendix-Down), respectively.

## References

[1]  Trevor Hastie, Robert Tibshirani, Balasubramanian Narasimhan and Gilbert Chu (2018). impute: impute: Imputation for
  microarray data. R package version 1.56.0.
  
[2] Cosmin Lazar (2015). imputeLCMD: A collection of methods for left-censored missing data imputation. R package version 2.0.
  https://CRAN.R-project.org/package=imputeLCMD

## Appendix 

### Enriched pathways or gene sets (higher) {#appendix-Up}

```{r Upregul, echo=F }
library("kableExtra")
kable(params$UpRegul, caption = " ") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 
```

###  Enriched pathways or gene sets (lower) {#appendix-Down}

```{r DoRegul, echo=F }
library("kableExtra")
kable(params$DoRegul) %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 
```



